\\fontsize{12pt}{3pt}\\selectfont \\textbf{\\textit{Journal: Ecosphere. Manuscript ID: ECS23-0342. Manuscript type: Article}}

### 2023-10-31 manuscript/manuscriptRW.Rmd
change supplementary wiyh appendix

The simulations were run in the absence of natural disturbances. Only
the model LPJ-GUESS had to include a "background" disturbance
probability to increase chances of shade-intolerant species to
establish. The simulations were set up to sample species-specific
recruitment rates per decade and per ha in the equilibrium state of the
model, typically entailing a "spin-up" run (as we did not provide any
forest data). The modeling teams decided on the simulated area and how
they derived these samples. The exact length of the simulation was also
decided by the modeling teams (Table \@ref(tab:tableModels)). Further
details on how each modeling team prepared the simulations and the
outputs are available in Supplementary Material 1.

@@

```{=tex}
\begin{equation}
Reineke = \frac{N7}{N10}^{-1.605} \label{eqReineke}
\end{equation}
```
We assessed model performance in relation to model traits focusing on
(1) model complexity as defined by @bugmann2022, (2) model type
(empirical or process-based), (3) the presence or absence of a canopy
feedback for regeneration, and (4) the scale of application of the model
(stand, landscape or global). We tested for significant differences
between simulated and observed values of the recruitment levels and
species diversity by using t-tests with a correction for multiple
testing. Note that the t-tests does not present conclusive evidence for
a specific hypotheses, but rather facilitates managing the extensive
number of comparisons discussed in our study.

@@

In most models, there were only small but significant differences in the
species diversity of recruitment between 7 and 10 cm across sites
(Tables S2). Four models (iLand, Landis-II, TreeMig and LPJ-GUESS)
maintained the recruitment diversity between the 7 and 10 cm thresholds
(i.e., the differences between them were not significant, see Table S2),
and the same was evident from the observed data.

For both the observed and the simulated data, and both recruitment
thresholds, species diversity in the recruitment compared to stand-level
diversity did not feature clear patterns across the models (Figures S1
and S2). Most models captured reasonably well or overpredicted species
diversity at the stand level (Figure 4A, left and center; cf. Figure
S3), but a characteristic feature was that the simulations had a much
lower variability of diversity compared to observations. Models
overpredicting stand-level species diversity included ForClim1, ForClim
11, FORMIND and TreeMig (Figure 4A, left). Several models underpredicted
species diversity, i.e., LPJ-GUESS, ForCEEPS(f) and 4C (Figure 4A,
right).

The majority of the models overpredicted recruitment diversity; as
already observed in Figure 3, only two of them underestimated it (4C and
Landis-II) (Figure 4B and Figure S4). Both the models with and without
feedback from the adult trees to regeneration (via seed production)
overestimated, underestimated or captured reasonably well recruitment
species diversity compared to the observations. Again, simulated
diversity had considerably lower variability than observed diversity in
the recruitment.

@@

There were strong differences among models regarding the mortality rate
between 7 and 10 cm (Figure 5). Most of the models featured a mortality
rate significantly larger than the observed data, thus at least
partially compensating for the general overestimation of recruitment
levels at 7 cm (Figure 2 and Table S3); six models did not feature a
significant difference compared to the observed data (ForCEEPS, FORMIND,
ForClim 1 and 11, iLand and aDGVM2). Four models (xComp, PICUS, LandClim
and TreeMig) featured very high mortality rates (i.e., well above the
Reineke self-thinning line), which compensated for the strong initial
overestimation of recruitment (cf. Figure 2). Interestingly, two models
that underestimated overall recruitment levels (SIBYLA and Landis-II;
cf. Figure 2) featured mortality rates that were close to but still
above the self-thinning line (Figure 5).

@@

Lastly, some models that had a low number of recruits at the 7 cm
threshold also had a small decrease of recruitment between 7 and 10 cm
(ratio 7-10 cm) (Figure S5). In these models, competition and
self-thinning are either not pronounced or must have occurred before the
trees had reached 7 cm. However, this pattern was not consistent across
models (Figure S5). It is noteworthy that there is a relationship
between the recruitment levels at 7 cm and the mortality rate (Table
S4): most of the models showed a positive effect (higher regeneration at
7 cm is coupled to an increase of the mortality between 7 and 10 cm).
While some models showed a negative effect, the linear trend was not
pronounced and not always significant regarding its slope.

@@

There was no significant relationship between the mean complexity of the
regeneration module (Table \@ref(tab:tableModels)) and the
overestimation proportion at a diameter threshold of 7 cm (Figure S6).
All models except 4C had significant differences between the observed
and simulated mean recruitment values. Therefore it was not possible to
differentiate in terms of performance between models type (empirical or
process based), or their scales (stand, landscape or global) (Table S5).
The values of species diversity in the recruitment simulated by each
model were also assessed against the observed data to see if there was a
difference between models that include a feedback compared to those that
did not; all the models except ForCEEPs(f) and FORMIND had significant
differences (Table S6). It is noteworthy that for the model ForCEEPS the
pattern changed from the overprediction of species diversity in the
regeneration to a diversity level that is closer to observations when
the recruitment module included feedback (ForCEEPs(f)).

@@

When evaluating total tree recruitment levels along key gradients of
light availability (basal area), temperature (degree-days) and soil
moisture (climatic water balance), distinct features emerged: the models
reproduced the effect of basal area in both its magnitude and patterns
(Figure 7 and S7) considerably better than the effects of the climatic
gradients, where they featured varying patterns.

@@

Silver fir (*Abies alba*), beech (*Fagus sylvatica*), Scots pine (*Pinus
sylvestris*) and oak (*Quercus* spp.) were the species observed most
frequently in the EuFoRIa data. Most models captured well the share of
basal area in the recruitment of these five species compared to the
observed data along the two major climatic gradients (Figure 8, cf.
Figure S8). However, the models differed considerably in the way in
which they simulated these climatic niches.

First, some models overestimated recruitment across the entire gradients
of at least one of the five main tree species (Figure 8, e.g., ForCEEPS,
ForCEEPS(f), LandClim or LPJ-GUESS), while other models overestimated
recruitment of more than one species for a large part of the climate
gradients, such as xComp, PICUS, SIBYLA or LPJ-GUESS (cf. red colors in
Figure 8). These trends were consistent for both recruitment thresholds,
although the exact changes in the share of the recruited basal area were
different (Figure S8).

@@

**Table 1.** Models included in this study, their characteristics and
simulation strategies used. Scale refers to the models scale, Type to
the models type, Population structure to the model population approach,
Feedback to the inclusion or not of adult tree feedback to the level and
species composition of regeneration, Approach to the model regeneration
module approach, Start from to the stage from which trees are recruited,
Complexity, to the mean regeneration formulation as estimated in
(Bugmann & Seidl, 2022), Species to which species from the original
eleven species were simulated, Runtime (spin up) [years] to the number
of years used per model in their simulations for the spin up period and
regeneration sampling; Climate data type (use of years) to the type of
data used in the simulations and how it was used, and Reference provides
the link to the publication with more model details. Further information
can be found on the individual model reports available in the
supplementary materials.

>>>>

The simulations were run in the absence of natural disturbances. Only
the model LPJ-GUESS had to include a "background" disturbance
probability to increase chances of shade-intolerant species to
establish. The simulations were set up to sample species-specific
recruitment rates per decade and per ha in the equilibrium state of the
model, typically entailing a "spin-up" run (as we did not provide any
forest data). The modeling teams decided on the simulated area and how
they derived these samples. The exact length of the simulation was also
decided by the modeling teams (Table \@ref(tab:tableModels)). Further
details on how each modeling team prepared the simulations and the
outputs are available in Appendix S1: Model reports.

@@

```{=tex}
\begin{equation}
Reineke = \frac{N7}{N10}^{-1.605} \label{eqReineke}
\end{equation}
```
We assessed model performance in relation to model traits focusing on
(1) model complexity as defined by @bugmann2022, (2) model type
(empirical or process-based), (3) the presence or absence of a canopy
feedback for regeneration, and (4) the scale of application of the model
(stand, landscape or global). We tested for significant differences
between simulated and observed values of the recruitment levels and
species diversity by using t-tests with the Bonferroni correction for
multiple testing. Note that the t-tests does not present conclusive
evidence for a specific hypotheses, but rather facilitates managing the
extensive number of comparisons discussed in our study.

@@

In most models, there were only small but significant differences in the
species diversity of recruitment between 7 and 10 cm across sites
(Appendix S1: Table S2). Four models (iLand, Landis-II, TreeMig and
LPJ-GUESS) maintained the recruitment diversity between the 7 and 10 cm
thresholds (i.e., the differences between them were not significant, see
Appendix S1: Table S2), and the same was evident from the observed data.

For both the observed and the simulated data, and both recruitment
thresholds, species diversity in the recruitment compared to stand-level
diversity did not feature clear patterns across the models (Appendix S1:
Figures S1 and S2). Most models captured reasonably well or
overpredicted species diversity at the stand level (Figure 4A, left and
center; cf. Appendix S1: Figure S3), but a characteristic feature was
that the simulations had a much lower variability of diversity compared
to observations. Models overpredicting stand-level species diversity
included ForClim1, ForClim 11, FORMIND and TreeMig (Figure 4A, left).
Several models underpredicted species diversity, i.e., LPJ-GUESS,
ForCEEPS(f) and 4C (Figure 4A, right).

The majority of the models overpredicted recruitment diversity; as
already observed in Figure 3, only two of them underestimated it (4C and
Landis-II) (Figure 4B and Appendix S1: Figure S4). Both the models with
and without feedback from the adult trees to regeneration (via seed
production) overestimated, underestimated or captured reasonably well
recruitment species diversity compared to the observations. Again,
simulated diversity had considerably lower variability than observed
diversity in the recruitment.

@@

There were strong differences among models regarding the mortality rate
between 7 and 10 cm (Figure 5). Most of the models featured a mortality
rate significantly larger than the observed data, thus at least
partially compensating for the general overestimation of recruitment
levels at 7 cm (Figure 2 and Appendix S1: Table S3); six models did not
feature a significant difference compared to the observed data
(ForCEEPS, FORMIND, ForClim 1 and 11, iLand and aDGVM2). Four models
(xComp, PICUS, LandClim and TreeMig) featured very high mortality rates
(i.e., well above the Reineke self-thinning line), which compensated for
the strong initial overestimation of recruitment (cf. Figure 2).
Interestingly, two models that underestimated overall recruitment levels
(SIBYLA and Landis-II; cf. Figure 2) featured mortality rates that were
close to but still above the self-thinning line (Figure 5).

@@

Lastly, some models that had a low number of recruits at the 7 cm
threshold also had a small decrease of recruitment between 7 and 10 cm
(ratio 7-10 cm) (Appendix S1: Figure S5). In these models, competition
and self-thinning are either not pronounced or must have occurred before
the trees had reached 7 cm. However, this pattern was not consistent
across models (Appendix S1: Figure S5). It is noteworthy that there is a
relationship between the recruitment levels at 7 cm and the mortality
rate (Appendix S1: Table S4): most of the models showed a positive
effect (higher regeneration at 7 cm is coupled to an increase of the
mortality between 7 and 10 cm). While some models showed a negative
effect, the linear trend was not pronounced and not always significant
regarding its slope.

@@

There was no significant relationship between the mean complexity of the
regeneration module (Table \@ref(tab:tableModels)) and the
overestimation proportion at a diameter threshold of 7 cm (Appendix S1:
Figure S6). All models except 4C had significant differences between the
observed and simulated mean recruitment values. Therefore it was not
possible to differentiate in terms of performance between models type
(empirical or process based), or their scales (stand, landscape or
global) (Appendix S1: Table S5). The values of species diversity in the
recruitment simulated by each model were also assessed against the
observed data to see if there was a difference between models that
include a feedback compared to those that did not; all the models except
ForCEEPs(f) and FORMIND had significant differences (Appendix S1: Table
S6). It is noteworthy that for the model ForCEEPS the pattern changed
from the overprediction of species diversity in the regeneration to a
diversity level that is closer to observations when the recruitment
module included feedback (ForCEEPs(f)).

@@

When evaluating total tree recruitment levels along key gradients of
light availability (basal area), temperature (degree-days) and soil
moisture (climatic water balance), distinct features emerged: the models
reproduced the effect of basal area in both its magnitude and patterns
(Figure 7 and Appendix S1: Figure S7) considerably better than the
effects of the climatic gradients, where they featured varying patterns.

@@

Silver fir (*Abies alba*), beech (*Fagus sylvatica*), Scots pine (*Pinus
sylvestris*) and oak (*Quercus* spp.) were the species observed most
frequently in the EuFoRIa data. Most models captured well the share of
basal area in the recruitment of these five species compared to the
observed data along the two major climatic gradients (Figure 8, cf.
Appendix S1: Figure S8). However, the models differed considerably in
the way in which they simulated these climatic niches.

First, some models overestimated recruitment across the entire gradients
of at least one of the five main tree species (Figure 8, e.g., ForCEEPS,
ForCEEPS(f), LandClim or LPJ-GUESS), while other models overestimated
recruitment of more than one species for a large part of the climate
gradients, such as xComp, PICUS, SIBYLA or LPJ-GUESS (cf. red colors in
Figure 8). These trends were consistent for both recruitment thresholds,
although the exact changes in the share of the recruited basal area were
different (Appendix S1: Figure S8).

@@

**Table 1.** Models included in this study, their characteristics and
simulation strategies used. Scale refers to the models scale, Type to
the models type, Population structure to the model population approach,
Feedback to the inclusion or not of adult tree feedback to the level and
species composition of regeneration, Approach to the model regeneration
module approach, Start from to the stage from which trees are recruited,
Complexity, to the mean regeneration formulation as estimated in
(Bugmann & Seidl, 2022), Species to which species from the original
eleven species were simulated, Runtime (spin up) [years] to the number
of years used per model in their simulations for the spin up period and
regeneration sampling; Climate data type (use of years) to the type of
data used in the simulations and how it was used, and Reference provides
the link to the publication with more model details. Further information
can be found on the individual model reports available in the Appendix
S1: Model reports.
